The Error Estimate of Force Calculation in the Inhomogeneous Molecular Systems

### 2011-10-18 shortRange.inhomo/writeDown/main.tex
backup the revise according to Luigi

\section{Introduction}
Non-bonded short-range interaction is encountered in nearly every
molecular simulation.  A naive idea to calculate short-range
interactions is to explicitly calculate all pairwise
interactions. This results in a computational cost scaling $\mathcal
O(N^2)$ per time step, which rapidly becomes inefficient as the number
of particles $N$ grows large. A better way is to introduce a cut-off
radius, outside of which all pairwise interactions are simply
neglected. In combination with the cell list and the neighbor list
algorithms \cite{frenkel02b}, the total computational cost of the
short-range interactions can be reduced to an acceptable level of
$\mathcal O(N)$.  The cut-off radius neglect all pairwise interactions
outside the cut-off radius, so it introduces error in the interaction
calculation. The cut-offed system could be viewed as approximation of
the original system, therefore, it is very important to test the
convergence of the properties under study with respect to the cut-off
radius before deriving conclusions on the origianl system.

Fortunately,
in most homogeneous systems, a satisfactory convergence can be
achieved even with a very small cut-off value, for example,
$2.5\sigma$.  Moreover, it is possible to correct the systematic error
of the potential energy, the pressure and the free energy by applying
the standard long range correction (LRC)~\cite{allen87a}, which
assumes the uniformity outside the cut-off radius and integrates the
contribution of the mentioned properties to infinity. 
% The spacial inhomogenety of the system in terms of particle density,
% particle type and change distribution is very common in natrual
% system.

The inhomogeneous systems, in which the density of certain type of
atom changes within the length scale of several molecules, are more
universal in nature. However, the direct molecular simulation of
inhomogeneous systems by the cut-off method will encounter non-trival
difficulties that do not exist in the homogeneous system simulations.
The simulation results are found to converge only at a very large
cut-off radius value~\cite{duque2004some, ismail2007application}, also
denpend on the energy/force continuity at the cut-off and the way of
applying the long-range corrections~\cite{trokhymchuk1999computer,
  guo1997long, mecke1997molecular, janecek2006long, goujon2004monte,
  shen2007comparative}.  Therefore, it is necessary to systemetically
study the precision of the short-range interaction calculation, and to
develop methods screening out the aforementioned cut-off artifect in
the inhomogeneous systems.

@@

In this work, we develop the error estimate of the short-range force
in inhomogeneous systems. We prove that the 
force error is comprised of a homogeneity error and a inhomogeneity
error, regardless the particle-particle correlation outside the
cut-off radius.  In a inhomogeneous system composed by several bulk
regions, an important observation is that the inhomogeneity error can
be more than one order of maganitude larger than the homogeneity
error, and plays a dominant role in the interfacial region. An
adaptive cut-off method is hereby proposed to equally distribute the
error across the simulation region, so that a substantial
computational effort can be saved while the precision of simulation is
preserved. We also proposed a long-range force correction (LFC) method
based on the error estimate to screen out the contribution of the
inhomogeneity error, which can tremendously improve the precision of
simulation without much extra computational load. To show the
effectiveness of the error estimate and the proposed methods, two
systems are tested: the liquid-vapor equilibrium for equilibrium
properties, and the colliding nanoscale clusters for dynamical
properties. Finally, conclusions are given in section
\ref{sec:conclusion}.

@@

\section{Theoretical background}
\subsection{Error estimate in a inhomogeneous system}

In this paper, we only consider the single-component system. Because
this assumption simplifies the discussion, and the error estimates for
the single-component system can be straightforwardly extended to the
multi-component system without any difficulty.  Only the
force error is estimated in the present paper, because the
applications are based on molecular dynamics simulation.
% Since the error estimate developed for the one-component system can be
% directly extended to multi-component system without substantial
% difficulty, the discussion in this paper will focus on one-component
% system for symplicity. 
Suppose the system is composed by $N$ identical particles located at
$\v r_1, \v r_2, \cdots, \v r_N$ with periodic boundary condition.
These particles are interacting via a short-range pairwise interaction
$u(\v r)$ with a cut-off radius $r_c > 0$. The corresponding force is
denoted by $\v f(\v r)$. 
% An infinite cut-off radius will lead to
% a precise calculation of interaction, but it is impractical.
% The error of computing interaction is mainly introduced by a finit
% value of the cut-off radius.
The force error mainly stems from the neglected interaction outside
the cut-off radius, so it is convenient to consider the complementary
of the cut-offed force that is defined by
\begin{align}
  \v f^c(\v r) =
  \left\{
  \begin{array}{ll}
    0, & \vert\v r\vert\leq r_c; \\
    \v f(\v r), & \vert\v r\vert > r_c.
  \end{array}
  \right.
\end{align}
We measure the force error at an arbitary position in the system,
saying $\v r$, by adding a testing particle to that position.  We
assume all particles in the system exert forces on the testng
particle, whereas the testing particle do NOT exert any force on the
system.  The exact force and the computed force (or cut-offed force)
exerted by the system on the testing particle are denoted by ${\v
  F}(\v r)$ and $\tilde{\v F}(\v r)$, respectively.  The difference
between them, namely the \emph{error force}, is denoted by $\Delta\v
F(\v r) = \v F(\v r) - \tilde{\v F}(\v r)$, which can be expressed by
\begin{align}
  \Delta\v F(\v r) = \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n).
\end{align}
Where $\v n$ is the box vector, so the summation over $\v n$ indicates
that all periodic images of the system are considered. In a
homogeneous system, the surounding of any particle is isotropic, so
the \emph{mean error force} vanishes: $\langle\Delta\v F(\v r)\rangle
= 0$. In contrast, the mean error force is not necessarily zero in a
inhomogeneous system:
\begin{align} \label{eqn:tmp3} %\nonumber
  \langle\Delta\v F(\v r)\rangle
  &=
  \big\langle
  \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n)
  \big\rangle 
  =
  \int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r',
\end{align}
where $\rho(\v r)$ denotes the particle number density at the position
$\v r$, and is periodically extended to $\mathbb R^3$. And notice the
error force implicitly depends on the cut-off radius used for the
interaction calculation.

The widely accepted definition of the force error is the \emph{root
  mean square} (RMS) \emph{error}, which is the square root of the second
moment of the error force $\mathcal E(\v r) =
\sqrt{\langle\vert\Delta\v F(\v r)\vert^2\rangle}$. It can be
calculate by
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  =\,&
  \big\langle\sum_{j,k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \big\langle\sum_j\vert\v f^c(\v r - \v r_j)\vert^2\big\rangle +
  \big\langle\sum_{j\neq k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r'
  + \\\label{eqn:tmp4}
  \,&
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\rho(\v r', \v r'')\,\d d\v r'\d d\v r''.
\end{align}
Where  $\rho(\v r', \v r'')$ is the pair density. By using the identity
\begin{align}  \label{eqn:tmp7}
  \rho(\v r', \v r'')
  = \rho(\v r')\rho(\v r'') + C (\v r', \v r''),
\end{align}
where 
\begin{align}
C (\v r', \v r'') = [\,\rho(\v r', \v r'') -  \rho(\v r')\rho(\v r'')\,]
\end{align}
denotes the density correlation between position $\v r'$ and $\v
r''$. Eqn. \eqref{eqn:tmp4} becomes
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  = &\,
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r' \,+ \\\nonumber
  &\,
  \bigg[\int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r'\,\bigg]^2 + \\\nonumber
  &\,
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\,C(\v r', \v r'')\,\d d\v r'\d d\v r'' \\\label{eqn:tmp5}
  = &\,
  \mathcal E^2_{\textrm{homo}}(\v r) +
%  \langle\Delta\v F(\v r)\,\rangle^2 +
  \mathcal E^2_{\textrm{inhomo}}(\v r) +
  \mathcal E^2_{\textrm{correlation}}(\v r).
\end{align}
If the system is homogeneous and the particle-particle correlation is
neglected, then only the first term is left on the right hand side of \eqref{eqn:tmp5}, so it
is called the homogeneity error.  This term originates from the
flucuation of the error force $\Delta\v F(\v r)$.  The second term is
called the inhomogeneity error, because it stems from the density
inhomogeneity of the system. It is worth poining out that the
inhomogeneity error is the nothing but the maganitude of the mean error force,
namely $\mathcal E_{\textrm{inhomo}}(\v r) = \vert\langle\Delta\v F(\v
r)\,\rangle\vert$.  The last term on right hand side of \eqref{eqn:tmp5} originates from the
correlation between the particles, so it is called the correlation
error.
% The correlation error involves a double integral, which
% implies that calculating this term requires more compuational cost
% than the homogeneity and inhomogeneity errors.
The rest of this paper
assumes the 
correlation error does not dominate in the system, and can be savely
neglected.

@@

The naive calculation of the convolutions in the error estimate \eqref{eqn:tmp5} requires
a compuational cost of $\mathcal O(N^2)$. To estimate the force error
on the fly, the compuational cost of the estimate should be reduced to
at least $\mathcal O(N\log N)$.  This is achieved by  using the fast Fourier transform. We
assume the simulation box is uniformly divided in to small bins.  The
positions of these bins are
\begin{align}
  \v r_{i_1, i_2, i_3} & =
  \frac{i_1}{K_1}\v a_1 + 
  \frac{i_2}{K_2}\v a_2 + 
  \frac{i_3}{K_3}\v a_3,\qquad 0\leq i_\alpha< K_\alpha
\end{align}
where $\v a_\alpha$ are box vectors, and $K_\alpha$ is the number of
division on each direction. In the reciprocal space, a corresponding
lattice is set up:
\begin{align}
  \v k_{k_1, k_2, k_3} & =
  k_1\v a_1^\ast +
  k_2\v a_2^\ast +
  k_3\v a_3^\ast, \qquad 0\leq k_\alpha<K_\alpha
\end{align}
where $\v a_\alpha^\ast$ are reciprocal box vectors defined by $\v
a_\alpha\cdot\v a_\beta^\ast = \delta_{\alpha\beta}$.  In each bin the
error and the particles number density are assumed to be constant,
then, a Fourier expansion can be obtained:
\begin{align}\nonumber
  \mathcal E^2_{\textrm{homo}}(\v r_{i_1,i_2,i_3})
  & =
  \frac1V\sum_{k_1=-\infty}^\infty\sum_{k_2=-\infty}^\infty\sum_{k_3=-\infty}^\infty
  [{\mathcal E}^2_{\textrm{homo}}]^{\wedge}(\v k_{k_1,k_2,k_3})
  % e^{2\pi i\v k_{k_1,k_2,k_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg] \\\label{eqn:tmp13}
  & =
  \frac1V\sum_{k_1=0}^{K_1}\sum_{k_2=0}^{K_2}\sum_{k_3=0}^{K_3}
  \hat{K}_{\textrm{homo}}(\v k'_{k_1,k_2,k_3})\:
  \hat{\rho}(\v k_{k_1,k_2,k_3})
  % e^{2\pi i\v k_{k_1,k_2,k_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg].
\end{align}
Here we denote the convolution kernel $\vert\v f^c(\v r)\vert^2$ by
$K_{\textrm{homo}}(\v r)$.  The second equation in \eqref{eqn:tmp13}
holds due to the fact: the Fourier transform of the convolution
$K_{\textrm{homo}}\ast\rho$ is equal to the multiplication of Fourier transforms
$\hat{K}_{\textrm{homo}}$ and $\hat\rho$, which are given by
\begin{align}\label{eqn:f-kernel-homo}
  \hat K_{\textrm{homo}}(\v k)
  &=
  \int_{\mathbb R^3}K_{\textrm{homo}}(\v r)e^{-2\pi i\v k\cdot\v r}\,\d d\v r,
\end{align}
and
\begin{align}
  \hat\rho(\v k_{k_1,k_2,k_3})
  & =
  \frac{V}{K_1K_2K_3}
  \sum_{i_1=0}^{K_1}\sum_{i_2=0}^{K_2}\sum_{i_2=0}^{K_2}
  \rho(\v r_{i_1,i_2,i_3})
  \exp\bigg[
  -2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg].
\end{align}
In Eqn. \eqref{eqn:tmp13}, the prime on $\v k_{k_1,k_2,k_3}$ means
that the periodic image $k_\alpha - K_\alpha$ should use instead of
$k_\alpha$ when $k_\alpha \geq K_\alpha/2$.  Most of the short-range
interactions are isotropic, i.e.  $K_{\textrm{homo}}(\v r) =
K_{\textrm{homo}}(r)$, then the three-dimensional integral in
Eqn. \eqref{eqn:f-kernel-homo} can be simplified to a one-dimentional
integral:
\begin{align}\label{eqn:kernel-inte}
  \hat{K}_{\textrm{homo}}(\v k)
  =
  % \int_{\mathbb R^3}K_{\textrm{homo}}(r)e^{-2\pi i\v k\cdot\v r}\d d\v r
  % =
  \int_{r_c}^\infty \frac{2 r}k\, K_{\textrm{homo}}(r) \sin(2\pi kr)\,\d dr,
\end{align}
which can be calculated by numerical
integration from $r_c$ to a large upper bound, saying $R_c$. Perfact
convergence can be achieved by using $R_c = 30\sigma$ throughout the
paper. Since integral \eqref{eqn:kernel-inte} is independent with the
density of the system, it can be calculated for only once and stored
for reference.

Similarly, by \eqref{eqn:tmp3}, the Fourier mode of the the mean error
force is
\begin{align}\label{eqn:tmp17}
  \langle\Delta\v F\rangle^\wedge(\v k) =
  \hat{\v f}^c(\v k)\,
  \hat\rho(\v k),
\end{align}
where the Fourier transform of the complementary force is given by
\begin{align}\label{eqn:tmp18}
  \hat{\v f}^c(\v k) 
  & = 
  2\pi\frac{\v k}k\, i\,
  \bigg\{
  r_c^2\,u(r_c)
  \bigg[\,
  \frac{2\cos(2\pi kr)}{2\pi kr}
  -\frac{2\sin(2\pi kr)}{(2\pi kr)^2}
  \,\bigg] 
  -
  \int_{r_c}^\infty 2r\,u(r)\sin(2\pi kr)\,\d d\v r
  \bigg\}.
\end{align}

@@

A basic observation in a inhomogeneous system is that the RMS force
error is not uniformly distributed. The difference between the maximum
and minimum error can be more than one order of
maganitude,
% in one of the cases we studied,
see Fig. \ref{fig:tmp2}. Traditionally, a very large cut-off radius is
used to reduce the maximum error in the system to reach convergent
simulation results~\cite{ismail2007application}. Therefore, a huge
amount of compuational effort is wasted in the low-error region. The
idea of the adaptive cut-off radius simulation is to use larger
cut-off radius for particles in the high-error region to control the
force error, while to use smaller cut-off radius for particles in the
low-error region to save compuational cost.  The effectiveness of this
method relies on a diliberated cut-off choice for each particle so
that the force error is uniformly distributed across the simulation
region.  Unfortunately, the compuational expense of a naive
implementation of this method is at least $\mathcal O(N^2\log N)$.
Becasue to determine the cut-off radius, we should estimate the RMS
force error at least once for each particle, and the computational
cost of each force error estimate is $\mathcal O(N \log N)$.

@@

Figure \ref{fig:tmp2} presents the real RMS force error (in green
square and log-scaled) in a system using a uniform cut-off raidus of $r_c^\ast =
7.5$. The number density distribution along $x$ direction is shown by
the solid red line for reference.  The error is comparatively constant
at the bulk liquid and gas region. In contrast, two paaks, which are
more than one order of maganitude larger than the force errors in bulk
regions, form at the interfacial region.  The estimated RMS force
error is presented by solid blue line.  The estimate follows the real
error very well at the interfacial region, but slightly over estimate
the error in the bulk regions.  Two components of the force error,
namely the homogeneity error and inhomogeneity error are presented in
Fig. \ref{fig:tmp2} by dashed and dotted blue lines, respectively.  In
the bulk regions, the dashed blue line overlaps with the solid blue
line, that means the homogeneity error dominates.  In the interfacial
region, in the contrary, the inhomogeneity error dominate, because the
dotted blue line overlaps with the solid one.

@@

% However, the maximum
% force error of the force correction method is reduced by about one
% order of maganitude.
% We also observe that the liquid densities are systemetically higher
% than the reference results. This fact will be discussed later by
% comparing with literature results.

>>>>

\section{Introduction}
Non-bonded short-range interaction is encountered in nearly every
molecular simulation.  A naive idea to calculate short-range
interactions is to explicitly calculate all pairwise
interactions. This results in a computational cost scaling $\mathcal
O(N^2)$ per time step, which becomes rapidly inefficient as the number
of particles $N$ grows. A better way is to introduce a cut-off
radius, outside of which all pairwise interactions are simply
neglected. In combination with the cell list and the neighbor list
algorithms \cite{frenkel02b}, the total computational cost of the
short-range interactions can be reduced to an acceptable level of
$\mathcal O(N)$.  The cut-off radius neglect all pairwise interactions
outside the cut-off radius, so it introduces an error in the interaction
calculation. Since the cut-offed system can be viewed as an approximation of
the original system, it is very important to test the
convergence of properties of interest as a function of the cut-off
radius.
% before deriving conclusions on the origianl system.

Fortunately,
in most homogeneous systems, a satisfactory convergence can be
achieved even with a very small cut-off value, for example,
$2.5\sigma$.  Moreover, it is possible to correct the systematic error
of the potential energy, the pressure and the free energy by applying
the standard long range correction (LRC)~\cite{allen87a}, which
assumes the uniformity outside the cut-off radius and integrates the
contribution of the mentioned properties to infinity. 
% The spacial inhomogenety of the system in terms of particle density,
% particle type and change distribution is very common in natrual
% system.
The inhomogeneous systems, in which the density of certain type of
atom changes within the length scale of several molecules, 
and thus lack of uniformity, 
the direct molecular simulation 
by the cut-off method will encounter non-trivial
difficulties.
The simulation results are found to converge only at a very large
cut-off radius value~\cite{duque2004some, ismail2007application}, also
depend on the energy/force continuity at the cut-off and the way of
applying the long-range corrections~\cite{trokhymchuk1999computer,
  guo1997long, mecke1997molecular, janecek2006long, goujon2004monte,
  shen2007comparative}.  Therefore, it is necessary to systematically
study the precision of the short-range interaction calculation, and to
develop methods for screening out the aforementioned cut-off artifact in
the inhomogeneous systems.

@@

In this work, we develop the error estimate of the short-range force
in inhomogeneous systems. We prove that the 
force error is comprised of a homogeneity error and an inhomogeneity
error, regardless of the particle-particle correlation outside the
cut-off radius.  In an inhomogeneous system composed by several bulk
regions, an important observation is that the inhomogeneity error can
be more than one order of magnitude larger than the homogeneity
error, and plays a dominant role in the interfacial region. An
adaptive cut-off method is hereby proposed to equally distribute the
error across the simulation region, so that a substantial
computational effort can be saved while the precision of simulation is
preserved. We also proposed a long-range force correction (LFC) method
based on the error estimate to screen out the contribution of the
inhomogeneity error, which can tremendously improve the precision of
simulation without much extra computational load. To show the
effectiveness of the error estimate and the proposed methods, two
systems are tested: the liquid-vapor equilibrium for equilibrium
properties, and the colliding nanoscale clusters for dynamical
properties. Finally, conclusions are given in section
\ref{sec:conclusion}.

@@

\section{Theoretical background}
\subsection{Error estimate in an inhomogeneous system}

In this paper, for simplicity,
we will consider one-component systems, but
% we only consider the single-component system. Because
% this assumption simplifies the discussion, and
the error estimates for
the one-component system can be straightforwardly extended to the
multi-component systems.  we consider
force error in the present paper, because the
applications are based on molecular dynamics simulation.
% Since the error estimate developed for the one-component system can be
% directly extended to multi-component system without substantial
% difficulty, the discussion in this paper will focus on one-component
% system for symplicity. 
Suppose the system is composed by $N$ identical particles located at
$\v r_1, \v r_2, \cdots, \v r_N$ with periodic boundary condition.
These particles are interacting via a short-range pairwise interaction
$u(\v r)$ with a cut-off radius $r_c > 0$. The corresponding force is
denoted by $\v f(\v r)$. 
% An infinite cut-off radius will lead to
% a precise calculation of interaction, but it is impractical.
% The error of computing interaction is mainly introduced by a finit
% value of the cut-off radius.
The force error mainly stems from the neglected interaction outside
the cut-off radius, so it is convenient to consider the complementary
of the cut-offed force that is defined by
\begin{align}
  \v f^c(\v r) =
  \left\{
  \begin{array}{ll}
    0, & \vert\v r\vert\leq r_c; \\
    \v f(\v r), & \vert\v r\vert > r_c.
  \end{array}
  \right.
\end{align}
We measure the force error at an arbitrary position in the system,
saying $\v r$, by adding a testing particle to that position.  We
assume all particles in the system exert forces on the testing
particle, whereas the testing particle does NOT exert any force on the
system.  The exact force and the computed force (or cut-offed force)
exerted by the system on the testing particle are denoted by ${\v
  F}(\v r)$ and $\tilde{\v F}(\v r)$, respectively.  The difference
between them, namely the \emph{error force}, is denoted by $\Delta\v
F(\v r) = \v F(\v r) - \tilde{\v F}(\v r)$, which can be expressed by
\begin{align}
  \Delta\v F(\v r) = \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n).
\end{align}
Where $\v n$ is the box vector, so the summation over $\v n$ indicates
that all periodic images of the system are considered. In a
homogeneous system, the surrounding of any particle is isotropic, so
the \emph{mean error force} vanishes: $\langle\Delta\v F(\v r)\rangle
= 0$. In contrast, the mean error force is not necessarily zero in an
inhomogeneous system:
\begin{align} \label{eqn:tmp3} %\nonumber
  \langle\Delta\v F(\v r)\rangle
  &=
  \big\langle
  \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n)
  \big\rangle 
  =
  \int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r',
\end{align}
where $\rho(\v r)$ denotes the particle number density at the position
$\v r$, and is periodically extended to $\mathbb R^3$. Notice that the
error force implicitly depends on the cut-off radius used for the
interaction calculation.

The widely accepted definition of the force error is the \emph{root
  mean square} (RMS) \emph{error}, which is the square root of the second
moment of the error force $\mathcal E(\v r) =
\sqrt{\langle\vert\Delta\v F(\v r)\vert^2\rangle}$. This is 
calculated as:
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  =\,&
  \big\langle\sum_{j,k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \big\langle\sum_j\vert\v f^c(\v r - \v r_j)\vert^2\big\rangle +
  \big\langle\sum_{j\neq k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r'
  + \\\label{eqn:tmp4}
  \,&
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\rho(\v r', \v r'')\,\d d\v r'\d d\v r''.
\end{align}
Where  $\rho(\v r', \v r'')$ is the pair density. By using the identity
\begin{align}  \label{eqn:tmp7}
  \rho(\v r', \v r'')
  = \rho(\v r')\rho(\v r'') + C (\v r', \v r''),
\end{align}
where 
\begin{align}
C (\v r', \v r'') = [\,\rho(\v r', \v r'') -  \rho(\v r')\rho(\v r'')\,]
\end{align}
denotes the density correlation between position $\v r'$ and $\v
r''$. Eqn. \eqref{eqn:tmp4} becomes
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  = &\,
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r' \,+ \\\nonumber
  &\,
  \bigg[\int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r'\,\bigg]^2 + \\\nonumber
  &\,
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\,C(\v r', \v r'')\,\d d\v r'\d d\v r'' \\\label{eqn:tmp5}
  = &\,
  \mathcal E^2_{\textrm{homo}}(\v r) +
%  \langle\Delta\v F(\v r)\,\rangle^2 +
  \mathcal E^2_{\textrm{inhomo}}(\v r) +
  \mathcal E^2_{\textrm{correlation}}(\v r).
\end{align}
If the system is homogeneous and the particle-particle correlation is
neglected, then only the first term is left on the right hand side of \eqref{eqn:tmp5}, so it
is called the homogeneity error.  This term originates from the
fluctuation of the error force $\Delta\v F(\v r)$.  The second term is
called the inhomogeneity error, because it stems from the density
inhomogeneity of the system. It is worth pointing out that the
inhomogeneity error is nothing but the magnitude of the mean error force,
namely $\mathcal E_{\textrm{inhomo}}(\v r) = \vert\langle\Delta\v F(\v
r)\,\rangle\vert$.  The last term on right hand side of \eqref{eqn:tmp5} originates from the
correlation between the particles, so it is called the correlation
error.
% The correlation error involves a double integral, which
% implies that calculating this term requires more compuational cost
% than the homogeneity and inhomogeneity errors.
The rest of this paper
assumes the 
correlation error does not dominate in the system, and can be safely
neglected.

@@

The naive calculation of the convolutions in the error estimate \eqref{eqn:tmp5} requires
a computational cost of $\mathcal O(N^2)$. To estimate the force error
on the fly, the computational cost of the estimate should be reduced to
at least $\mathcal O(N\log N)$.  This is achieved by  using the fast Fourier transform. We
assume the simulation box is uniformly divided in to small bins.  The
positions of these bins are
\begin{align}
  \v r_{i_1, i_2, i_3} & =
  \frac{i_1}{K_1}\v a_1 + 
  \frac{i_2}{K_2}\v a_2 + 
  \frac{i_3}{K_3}\v a_3,\qquad 0\leq i_\alpha< K_\alpha
\end{align}
where $\v a_\alpha$ are box vectors, and $K_\alpha$ is the number of
division on each direction. In the reciprocal space, a corresponding
lattice is set up:
\begin{align}
  \v k_{k_1, k_2, k_3} & =
  k_1\v a_1^\ast +
  k_2\v a_2^\ast +
  k_3\v a_3^\ast, \qquad 0\leq k_\alpha<K_\alpha
\end{align}
where $\v a_\alpha^\ast$ are reciprocal box vectors defined by $\v
a_\alpha\cdot\v a_\beta^\ast = \delta_{\alpha\beta}$.  In each bin the
error and the particles number density are assumed to be constant,
then, a Fourier expansion can be obtained:
\begin{align}\nonumber
  \mathcal E^2_{\textrm{homo}}(\v r_{i_1,i_2,i_3})
  & =
  \frac1V\sum_{k_1=-\infty}^\infty\sum_{k_2=-\infty}^\infty\sum_{k_3=-\infty}^\infty
  [{\mathcal E}^2_{\textrm{homo}}]^{\wedge}(\v k_{k_1,k_2,k_3})
  % e^{2\pi i\v k_{k_1,k_2,k_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg] \\\label{eqn:tmp13}
  & =
  \frac1V\sum_{k_1=0}^{K_1}\sum_{k_2=0}^{K_2}\sum_{k_3=0}^{K_3}
  \hat{K}_{\textrm{homo}}(\v k'_{k_1,k_2,k_3})\:
  \hat{\rho}(\v k_{k_1,k_2,k_3})
  % e^{2\pi i\v k_{k_1,k_2,k_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg].
\end{align}
Here we denote the convolution kernel $\vert\v f^c(\v r)\vert^2$ by
$K_{\textrm{homo}}(\v r)$.  The second equation in \eqref{eqn:tmp13}
holds due to the fact: the Fourier transform of the convolution
$K_{\textrm{homo}}\ast\rho$ is equal to the multiplication of Fourier transforms
$\hat{K}_{\textrm{homo}}$ and $\hat\rho$, which are given by
\begin{align}\label{eqn:f-kernel-homo}
  \hat K_{\textrm{homo}}(\v k)
  &=
  \int_{\mathbb R^3}K_{\textrm{homo}}(\v r)e^{-2\pi i\v k\cdot\v r}\,\d d\v r,
\end{align}
and
\begin{align}
  \hat\rho(\v k_{k_1,k_2,k_3})
  & =
  \frac{V}{K_1K_2K_3}
  \sum_{i_1=0}^{K_1}\sum_{i_2=0}^{K_2}\sum_{i_2=0}^{K_2}
  \rho(\v r_{i_1,i_2,i_3})
  \exp\bigg[
  -2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg].
\end{align}
Note that in Eqn. \eqref{eqn:tmp13}, the prime on $\v k_{k_1,k_2,k_3}$ means
that the periodic image $k_\alpha - K_\alpha$ should use instead of
$k_\alpha$ when $k_\alpha \geq K_\alpha/2$.
In Eqn. \eqref{eqn:f-kernel-homo}
the Fourier transform is performed on $K_{\textrm{homo}}(\v r)$ rather than
$K_{\textrm{homo}}(\v r - \v r')$.
Most of the short-range
interactions are isotropic, so  $K_{\textrm{homo}}(\v r) =
K_{\textrm{homo}}(r)$, then the three-dimensional integral in
Eqn. \eqref{eqn:f-kernel-homo} can be simplified to a one-dimensional
integral:
\begin{align}\label{eqn:kernel-inte}
  \hat{K}_{\textrm{homo}}(\v k)
  =
  % \int_{\mathbb R^3}K_{\textrm{homo}}(r)e^{-2\pi i\v k\cdot\v r}\d d\v r
  % =
  \int_{r_c}^\infty \frac{2 r}k\, K_{\textrm{homo}}(r) \sin(2\pi kr)\,\d dr,
\end{align}
which can be calculated by numerical
integration from $r_c$ to a large upper bound, saying $R_c$. Perfect
convergence can be achieved by using $R_c = 30\sigma$ throughout the
paper. Since integral \eqref{eqn:kernel-inte} is independent of the
density of the system, it is calculated only once and stored.

Similarly, for Eqn. \eqref{eqn:tmp3}, the Fourier mode of the the mean error
force is
\begin{align}\label{eqn:tmp17}
  \langle\Delta\v F\rangle^\wedge(\v k) =
  \hat{\v f}^c(\v k)\,
  \hat\rho(\v k),
\end{align}
where the Fourier transform of the complementary force is given by
\begin{align}\label{eqn:tmp18}
  \hat{\v f}^c(\v k) 
  & = 
  2\pi\frac{\v k}k\, i\,
  \bigg\{
  r_c^2\,u(r_c)
  \bigg[\,
  \frac{2\cos(2\pi kr)}{2\pi kr}
  -\frac{2\sin(2\pi kr)}{(2\pi kr)^2}
  \,\bigg] 
  -
  \int_{r_c}^\infty 2r\,u(r)\sin(2\pi kr)\,\d d\v r
  \bigg\}.
\end{align}

@@

A basic observation in an inhomogeneous system is that the RMS force
error is not uniformly distributed. The difference between the maximum
and minimum error can be more than one order of
magnitude
% in one of the cases we studied,
(see Fig. \ref{fig:tmp2}). Traditionally, a very large cut-off radius is
used to reduce the maximum error in the system to reach convergent
simulation results~\cite{ismail2007application}. Therefore, a huge
amount of computational effort is wasted in the low-error region. The
idea of the adaptive cut-off radius simulation is to use larger
cut-off radius for particles in the high-error region to control the
force error, while to use smaller cut-off radius for particles in the
low-error region to save computational cost.  The effectiveness of this
method relies on a deliberated cut-off choice for each particle so
that the force error is uniformly distributed across the simulation
region.  Unfortunately, the computational expense of a naive
implementation of this method is at least $\mathcal O(N^2\log N)$.
Because in order to determine the cut-off radius, we should estimate the RMS
force error at least once for each particle, and the computational
cost of each force error estimate is $\mathcal O(N \log N)$.

@@

Fig.~\ref{fig:tmp2} presents the real RMS force error (in green
square and log-scaled) in a system using a uniform cut-off radius of $r_c^\ast =
7.5$. The number density distribution along $x$ direction is shown by
the solid red line for reference.  The error is comparatively constant
at the bulk liquid and gas region. In contrast, two peaks, which are
more than one order of magnitude larger than the force errors in bulk
regions, form at the interfacial region.  The estimated RMS force
error is presented by solid blue line.  The estimate follows the real
error very well at the interfacial region, but slightly over estimate
the error in the bulk regions.  Two components of the force error,
namely the homogeneity error and inhomogeneity error are presented in
Fig.~\ref{fig:tmp2} by dashed and dotted blue lines, respectively.  In
the bulk regions, the dashed blue line overlaps with the solid blue
line, that means the homogeneity error dominates.  In the interfacial
region, instead, the inhomogeneity error dominate, because the
dotted blue line overlaps with the solid one.

@@

% However, the maximum
% force error of the force correction method is reduced by about one
% order of maganitude.
% We also observe that the liquid densities are systemetically higher
% than the reference results. This fact will be discussed latter by
% comparing with literature results.
